Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-04-08 manuscript.qmd
define scenarios, write methods, add upsetplot

## Parameter Symmetries

@@

## Tumors Extending Over the Mid-Sagittal Line

@@

@@

>>>>

## Parameter Symmetries {#sec-params-symmetry}

@@

## Tumors Extending Over the Mid-Sagittal Line {#sec-midline}

@@

- describe origin of data
- why is this data special?
- mention [LyProX](https://lyprox.org)
- visualize data, e.g. using upsetplots

@@

# Methods

We compared three different aprroaches to modelling the contralateral involvement in OPSCC:

1. A naive model of the contralateral side only. Neither the patterns of ipsilateral involvement, nor the information on whether the patients' tumors crossed the mid-sagittal line enter this model.
2. The bilateral model as introduced in the sections up to and including @sec-params-symmetry. The involvement patterns of both sides enter this model, but not whether the tumors extended the midline.
3. An additional bilateral model that tracks the midline extension of the patients' tumors as a random variable. This includes the formalism introduced up to and including @sec-midline. In contrast to the other two models, it does consider the reported midline extension in the data to attempt a more accurate prediction of the risk for contralateral occult disease.

- all models initialized with the same settings
  - same graph
  - same data
  - same distributions over diagnose time
- parameter samples were drawn for all three models:
  - the sampling used {{< var sampling.walkers_per_dim >}} walkers per parameter space dimension.
  - Every {{< var sampling.check_interval >}} steps, convergence was checked
  - convergence was reached when two criteria were met:
      1. the relative difference between subsequent autocorrelation estimates of the chain does not change more than {{< var sampling.rel_thresh |  >}}.
      2. the autocorrelation estimate of the chain drops below the length of the chain divided by {{< var sampling.trust_fac >}}.
  - after convergence, {{< var sampling.nsteps >}} samples (from the parallel walkers) were drawn that were spaced {{< var sampling.thin >}} steps apart, resulting in `num params x {{< var sampling.walkers_per_dim >}} x {{< var sampling.nsteps >}}` parameter samples.

For every model, we compare the predicted prevalence of a certain scenario to how often this scenario is seen in the data. The considered scenarios are choosen to test if the models can take into account T-category, the correlation of ipsi- and contralateral involvement, and the primary tumor's lateralization (whether or not it extends over the mid-sagittal line).

- [ ] list and describe the scenarios
